The Effect of Events on Discussion in the Australian Federal Parliament (1901--2017)

### 2018-10-22 outputs/docs/events/paper.Rmd
Start slides and update paper

We find: 1) changes in government tend be associated with topic changes even when the party in power does not change. For instance, the change from Hughes to Bruce in 1922, Menzies to Holt in 1966; Hawke to Keating in 1991; and Rudd to Gillard in 2010 are all associated with significant changes in topics despite no change in the party in power. 2) As expected, elections where the party in power also changes, such as Menzies in 1949, Whitlam in 1972, Fraser in 1975, Hawke in 1983, Howard in 1996, Rudd in 2007, and Abbott in 2013 are associated with topic changes, but the 1974 Whitlam, 1980 Fraser, and the 1998 and 2004 Howard re-elections, stand out as elections where the government was returned yet there was a significant change in topics. 3) Economic events, such as financial crises, have less significant effects than other events such as terrorism.

@@

# Data
## Parliamentary text data
Following the example of the UK a daily text record called Hansard of what was said in the Australian Federal Parliament has been made available since it was established in 1901.[^Verbatimfootnote] Earlier work on the influence of parliaments, such as @ZandenBuringhBosker2012, often examined broader activity measures such as counts of sitting days. This allowed for long time-frames and wide comparisons. But analysing Hansard records and their equivalents directly is increasingly viable as new methods and reduced computational costs make it easier.

@@

The topics that we have focused on here are:
12: war, defence, country, ill, soldiers; 
13: industry, trade, countries, development, country; 
17: security, time, world, war, country; 
22: war, country, matter, production, time; 
23: defence, time, national, service, vietnam; 
24: time, budget, tax, billion, million; 
29: education, schools, students, school, university; 
30: oil, energy, industry, nuclear, research; 
32: industry, wheat, wool, growers, farmers; 
33: per_cent, budget, time, tax, business; 
35: community, time, health, support, local; and
39: health, services, funding, care, million.

@@

We summarise the days where parliamentary discussion was significantly different from the rest prevailing in that week in Table \@ref(tab:keyeventsresultstable). **[THIS IS A CUT-DOWN LIST FOR DRAFT PURPOSES - REPLACE WITH ONE FROM IMPROVED MODEL.]**

@@

Of the 36 different governments over this period, we find that 12 **[UPDATE BASED ON NEW MODEL]** of them are significantly different to the government that preceded them. However, two of these results -- the significance of the Page  Government and the first Menzies Government -- are likely due to the short length of the Page Government and should be ignored, leaving only ten **[UPDATE BASED ON NEW MODEL]**. The earliest of these is the second Menzies government starting in 1949, and these ten are concentrated in the second half of our sample, with three of them being in the past twenty years **[UPDATE BASED ON NEW MODEL]**. Similarly, of the 45 general elections that have been held we find that 12 **[UPDATE BASED ON NEW MODEL]** of them define periods that were significantly different to their immediate predecessor. 1974, 1980, 1990, 1998, 2004, and 2007 stand out as elections where the government did not change, but the model suggests there was considerable change in the topics discussed in parliament **[UPDATE BASED ON NEW MODEL]**.

@@

The Menzies Government was succeeded by the Holt Government in January 1966. This is an example where there was a change in government without an election, as the next election only happened in November 1966. We find that the Holt Government is significantly different to the Menzies Government. In Figures \@ref(fig:mugovmeziesholt) we compare the topics during the final term of the Menzies Government with the topics of the Holt Government.

```{r mugovmeziesholt, cache = TRUE, echo = FALSE, fig.cap = "Government level effects on selected topics", out.width = '100%'}
knitr::include_graphics("../../figures/mu_election-finalmenzies-firstholt.pdf")
```

The Whitlam government is interesting as we find a difference in the topics after it was first elected in December 1972, compared with its second election win in May 1974. Figure \@ref(fig:muelectionwhitlam) compares the topics that are significant in the first Whitlam term and then compares them to those in the second Whitlam government.

```{r muelectionwhitlam, cache = TRUE, echo = FALSE, fig.cap = "Government level effects on selected topics", out.width = '100%'}
knitr::include_graphics("../../figures/mu_election-whitlamelections.pdf")
```

The Howard Government is interesting because of the significant differences between elections. For instance, each of the election periods is associated with fairly substantial differences compared with the preceding election periods, and all apart from 2001 are actually significantly different at the 95 per cent level. Figure \@ref(fig:muelectionhoward) compares the topics that are significant in the different Howard terms.

```{r muelectionhoward, cache = TRUE, echo = FALSE, fig.cap = "Different ", out.width = '100%'}
knitr::include_graphics("../../figures/mu_election-howard.pdf")
```

To a certain extent the change after the November 2001 election is expected because of the 9/11 terrorist attacks that had only occurred two months earlier, the Bali Bombings that occurred in October 2002, and the dramatic increase in the discussion related to terrorism and conflict over these years. However the change in 1998 and 2004 is more unexpected. Although the Howard Government is the second-longest serving government and commonly thought of as a period of stability because the senior ministers were consistent as well, it might be that it is better to think of the Howard Government as a combination of three or four different periods and that the Howard Government reinvented itself over this period. **[I WONDER IF THERE'S SOME REFERENCE THAT MAKES SIMILAR POINTS TO THIS?]**

**[Talk about the other significant results.]**

@@

The model identifies 11 September 2001, 12 October 2002.... **[Add more here.]**

>>>>

We find: 1) changes in government tend be associated with topic changes even when the party in power does not change. For instance, the change from Hughes to Bruce in 1922, Menzies to Holt in 1966; Hawke to Keating in 1991; and Rudd to Gillard in 2010 are all associated with significant changes in topics despite no change in the party in power. 2) As expected, elections where the party in power also changes, such as Fisher in 1910 and 1914, Menzies in 1949, Whitlam in 1972, Fraser in 1975, Hawke in 1983, Howard in 1996, Rudd in 2007, and Abbott in 2013 are associated with topic changes, but the 1974 Whitlam, 1980 Fraser, and the 1998 and 2004 Howard re-elections, stand out as elections where the government was returned yet there was a significant change in topics. 3) Economic events, such as financial crises, have less significant effects than other events such as terrorism.

@@

# Data

Following the example of the UK a daily text record called Hansard of what was said in the Australian Federal Parliament has been made available since it was established in 1901.[^Verbatimfootnote] Earlier work on the influence of parliaments, such as @ZandenBuringhBosker2012, often examined broader activity measures such as counts of sitting days. This allowed for long time-frames and wide comparisons. But analysing Hansard records and their equivalents directly is increasingly viable as new methods and reduced computational costs make it easier.

@@

The topics that we have focused on here are:
12, 17, 22, 23 which have to do with conflict, defence and security; 13 and 32 which have to do with commerce and trade; 24 and 33 which have to do with the budget, tax and spending; 29 which is to do with education and 39 which has to do with health.

@@

We summarise the days where parliamentary discussion was significantly different from the rest prevailing in that week in Table \@ref(tab:keyeventsresultstable).

@@

Of the 36 different governments over this period, we find that 14 of them are significantly different to the government that preceded them. However, two of these results -- the significance of the Page  Government and the first Menzies Government -- are likely due to the short length of the Page Government and should be ignored, leaving only 12. The earliest of these are the Second and Third Fisher Governments, which were different to the Third Deakin Government and the Cook Government respectively. To a certain extent this is likely tied up with changes due to World War I.

The Second Menzies Government, starting in 1949, is the next government that is significantly different to a precedessor. The rest of the ones that are different are concentrated in the second half of our sample, with three of them being in the past twenty years. Similarly, of the 45 general elections that have been held we find that 14 of them define periods that were significantly different to their immediate predecessor. 1974, 1980, 1990, 1998, 2004, and 2007 stand out as elections where the government did not change, but the model suggests there was considerable change in the topics discussed in parliament.

@@

The Menzies Government was succeeded by the Holt Government in January 1966. This is an example where there was a change in government without an election, as the next election only happened in November 1966. We find that the Holt Government is significantly different to the Menzies Government. In Figures \@ref(fig:muelectionall) we compare the topics during the final term of the Menzies Government with the topics of the Holt Government.

```{r muelectionall, cache = TRUE, echo = FALSE, fig.cap = "Differences between various elections", out.width = '100%'}
knitr::include_graphics("../../figures/mu_election-differences.pdf")
```

The Whitlam government is interesting as we find a difference in the topics after it was first elected in December 1972, compared with its second election win in May 1974. Figure \@ref(fig:muelectionall) compares the topics that are significant in the first Whitlam term and then compares them to those in the second Whitlam government.

The Howard Government is interesting because of the significant differences between elections. For instance, each of the election periods is associated with fairly substantial differences compared with the preceding election periods, and all apart from 2001 are actually significantly different at the 95 per cent level. Figure \@ref(fig:muelectionall) compares the topics that are significant in the different Howard terms.

To a certain extent the change after the November 2001 election is expected because of the 9/11 terrorist attacks that had only occurred two months earlier, the Bali Bombings that occurred in October 2002, and the dramatic increase in the discussion related to terrorism and conflict over these years. However the change in 1998 and 2004 is more unexpected. Although the Howard Government is the second-longest serving government and commonly thought of as a period of stability because the senior ministers were consistent as well, it might be that it is better to think of the Howard Government as a combination of three or four different periods and that the Howard Government reinvented itself over this period.

@@

The model identifies 11 September 2001, 12 October 2002.... **[It doesn't anymore. Not really sure what to do. here.]**
